---
title: "Analysis Code"
author: "Lorenzo Ricolfi"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    toc_depth: 3
    number_sections: no
    theme: cerulean
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

The following packages are required to run the code in this script.
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
#devtools::install_github("EIvimeyCook/ShinyDigitise")
#library(shinyDigitise)
pacman::p_load(tidyr,
               dplyr,
               here,
               ggplot2,
               gridExtra,
               metafor,
               orchaRd
               )       
```

```{r}
ma_details <- read.csv(here("data/ma_details.csv"))
ma_e_details <- read.csv(here("data/ma_e_details.csv"))
```

# Data wrangling and cleaning
```{r}
ma_e_details <- ma_e_details %>%
  filter(chemical_class == "PFAS", 
         !is.na(health_outcome_type),
         ma_es_measure != "unclear") %>%
  mutate(ma_es_measure = recode(
    ma_es_measure,
    "? (regression coefficient) value" = "Beta (regression coefficient) value",
    "beta (regression coefficient) value" = "Beta (regression coefficient) value",
    "Adjusted ? (regression coefficient)" = "Beta (regression coefficient) value",
    "adjusted coefficient regression (?)" = "Beta (regression coefficient) value",
    "relative risk" = "Risk ratio",
    "RR (risk ratio)" = "Risk ratio",
    "OR (odds ratio)" = "Odds ratio",
    "fishers z" = "Fisher's Z",
    "log OR (odds ratio)" = "log Odds ratio")) %>% 
  mutate(chemical_id = recode(
    chemical_id,
    "n-PFOA" = "PFOA",
    "n-PFOS" = "PFOS",
    "PFHsX" = "PFHxS")) %>% 
  left_join(ma_details, by = "ma_id")
```

# Exploratory analyses
## Effect size types
```{r}
es_count <- ma_e_details %>% 
  filter(!is.na(ma_es_measure) & ma_es_measure != "") %>%
  group_by(ma_es_measure) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_chemical_groups = n_distinct(chemical_class),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_health_groups = n_distinct(health_outcome_group),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>% 
  head(20)
# Convert the table to a grid for plotting
es_table <- tableGrob(es_count)
```

```{r, eval=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/es_table.png"), es_table, width = 12, height = 7)
```

## Chemical ID
```{r}
# Count the number of occurrences of each chemical_id
chemical_count <- ma_e_details %>%
  filter(!is.na(chemical_id) & chemical_id != "") %>% 
  group_by(chemical_id) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_health_outcome_group = n_distinct(health_outcome_group),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(20)
# Convert the table to a grid for plotting
chem_table <- tableGrob(chemical_count)
```

```{r, eval=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/chemicals_table.png"), chem_table, width = 11, height = 7)
```

## Health outcomes

```{r}
# Fix encoding issues in health_outcome_type
ma_e_details$health_outcome_type <- iconv(ma_e_details$health_outcome_type, from = "UTF-8", to = "UTF-8", sub = "")

# Count the number of occurrences of each health_outcome_type
outcome_type_table <- ma_e_details %>%
  filter(!is.na(health_outcome_type) & health_outcome_type != "") %>% 
  group_by(health_outcome_type) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate)),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(20)

# Convert the table to a grid for plotting
ho_table <- tableGrob(outcome_type_table)
```


```{r, eval=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/health_outcome_type_table.png"), ho_table, width = 11, height = 6)
```
## Health outcome group
```{r}
# Count the number of occurrences of each health_outcome_info
outcome_group_table <- ma_e_details %>%
  filter(!is.na(health_outcome_group) & health_outcome_group != "") %>% 
  group_by(health_outcome_group) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcome_types = n_distinct(health_outcome_type),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(8)

# Convert the table to a grid for plotting
ho_group_table <- tableGrob(outcome_group_table)
```


```{r, eval=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/health_outcome_group_table.png"), ho_group_table, width = 12, height = 5)
```

# Effect size conversion

See: Borenstein, M., Hedges, L. V., Higgins, J. P., & Rothstein, H. R. (2021). Introduction to meta-analysis. John wiley & sons.

```{r}
# Define a conversion function with an optional baseline risk parameter
convert_to_logOR <- function(effect_measure, point_estimate, ma_l_ci, ma_u_ci, regression_type = NA, baseline_risk = NA) {
  # Ensure the effect measure is treated as character
  effect_measure <- as.character(effect_measure)
  # Ensure point estimate, lower CI, and upper CI are numeric
  point_estimate <- suppressWarnings(as.numeric(as.character(point_estimate)))
  ma_l_ci <- suppressWarnings(as.numeric(as.character(ma_l_ci)))
  ma_u_ci <- suppressWarnings(as.numeric(as.character(ma_u_ci)))
  
  # Conversion for Odds ratio and log Odds ratio
  if (effect_measure == "Odds ratio") {
    # Convert OR to logOR
    return(c(log(point_estimate), log(ma_l_ci), log(ma_u_ci)))
    
  } else if (effect_measure == "log Odds ratio") {
    # Already in logOR
    return(c(point_estimate, ma_l_ci, ma_u_ci))
    
    # Conversion for SMD (standardized mean difference)
  } else if (effect_measure == "SMD (standardized mean difference)") {
    # Following Borenstein et al. (2021): logOR = SMD * pi/sqrt(3)
    logOR_point_estimate <- point_estimate * (pi / sqrt(3))
  
    # Convert lower and upper confidence intervals using the same logic
    logOR_l_ci <- ma_l_ci * (pi / sqrt(3))
    logOR_u_ci <- ma_u_ci * (pi / sqrt(3))
  
    return(c(logOR_point_estimate, logOR_l_ci, logOR_u_ci))

    # Conversion for Relative risk / Risk ratio with optional baseline risk
  } else if (effect_measure == "Risk ratio") {
    if (!is.na(baseline_risk)) {
      # Using the conversion: OR = RR * (1-p0) / (1-RR*p0)
      OR <- point_estimate * (1 - baseline_risk) / (1 - point_estimate * baseline_risk)
      OR_l_ci <- ma_l_ci * (1 - baseline_risk) / (1 - ma_l_ci * baseline_risk)
      OR_u_ci <- ma_u_ci * (1 - baseline_risk) / (1 - ma_u_ci * baseline_risk)
      return(c(log(OR), log(OR_l_ci), log(OR_u_ci)))
    } else {
      # If no baseline risk is provided, fallback to the approximation:
      return(c(log(point_estimate), log(ma_l_ci), log(ma_u_ci)))
    }
    
    # Conversion for Fisher's z (often used for correlations)
  } else if (grepl("fishers z", effect_measure, ignore.case = TRUE)) {
    # Convert z to correlation, then to logOR.
    # r = tanh(z)
    # Approximation: logOR = r * pi/sqrt(3)
    r <- tanh(point_estimate)
    return(r * (pi / sqrt(3)))
    
    # Conversion for regression coefficients assumed from logistic regression:
  } else if (effect_measure %in% c("Beta (regression coefficient) value")) {
    # Check if regression_type is 'logistic'
    if (!is.na(regression_type) && regression_type == "logistic") {
        return(c(point_estimate, ma_l_ci, ma_u_ci))
    } else {
        return(rep(NA, 3)) # Not logistic regression, cannot convert
    }
} else {
    warning("Effect measure conversion not implemented for: ", effect_measure)
    return(rep(NA, 3)) # Return NA for all three
  }
}
```

# Effect size calculation
```{r}
# Apply conversion to each row
ma_e_details <- ma_e_details %>%
  rowwise() %>%
  mutate(
    converted = list(convert_to_logOR(ma_es_measure, ma_point_estimate,
                                     ma_l_ci, ma_u_ci, regression_type, 0.1)),
    logOR = converted[1],
    l_ci_logOR = converted[2],
    u_ci_logOR = converted[3]
  ) %>%
  ungroup() %>%
  select(-converted) %>% 
  filter(!is.na(logOR))
```

# Model
```{r}
# Calculating effect size standard errors
# SE = (Upper bound – Lower bound) / (2 × 1.96)
ma_e_details <- ma_e_details %>%
  mutate(se = ((u_ci_logOR - l_ci_logOR)/(2*1.96)))  # Convert 95% CI to standard error (se)

# Fitting the multilevel meta-analytic model
mod <- rma.mv(yi = logOR,  # Effect size
              V = se^2, # Variance
              random = list(~1|ma_id, # Meta-analysis identity
                            ~1|ma_e_id), # Meta-analytic point estimate identity
              data = ma_e_details)

# Display model results
summary(mod)
# Total variance
tot_var <- sum(mod$sigma2)
# Between-study variance (tau^2_study)
mod$sigma2[1]
# Within-study variance (tau^2_effect)
mod$sigma2[2]
# Proportion of variance at each level
i2_study <- (mod$sigma2[1]/tot_var) * 100
i2_effect <- (mod$sigma2[2]/tot_var) * 100
cat("I² (between studies):", round(i2_study, 1), "%\n")
cat("I² (within studies):", round(i2_effect, 1), "%\n")
```

```{r}
i2_results <- i2_ml(mod)
# Print I² values
cat("I² (between studies):", round(i2_results["I2_ma_id"], 1), "%\n")
cat("I² (within studies):", round(i2_results["I2_ma_e_id"], 1), "%\n")
```


# Forest plot
```{r}
ma_e_details$ma_id <- factor(ma_e_details$ma_id, 
                            levels = c(levels(ma_e_details$ma_id), "Overall"))

ggplot(ma_e_details, aes(x = logOR, y = ma_id)) +
  geom_point(shape = 18, size = 4) +  # Point estimates
  geom_errorbarh(aes(xmin = l_ci_logOR, xmax = u_ci_logOR), 
                 height = 0.2) +      # Confidence intervals
  geom_vline(xintercept = 0,           # Line at null effect
             linetype = "dashed", 
             color = "red") +
  geom_point(data = data.frame(logOR = mod$b[1], 
                              ma_id = "Overall",
                              l_ci_logOR = mod$ci.lb,
                              u_ci_logOR = mod$ci.ub),
             shape = 23, size = 4, fill = "red") +
  geom_errorbarh(data = data.frame(logOR = mod$b[1], ma_id = "Overall"),
                 aes(xmin = mod$ci.lb, xmax = mod$ci.ub),
                 height = 0.3, size = 1) +
  labs(x = "log Odds Ratio", 
       y = "Meta-analysis",
       title = "Forest Plot of Effect Sizes") +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank()) + # Cleaner horizontal lines
  scale_y_discrete(limits = rev)
```

# OrchaRd
```{r}
# Create orchard plot with your specifications
orchard_plot <- orchard_plot(
  object = mod,
  group = "ma_id",
  xlab = "log Odds Ratio",
  transfm = "none"
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  )

# Display the plot
print(orchard_plot)
```
```{r}
orchard_plot2 <- orchard_plot(
  object = mod,
  group = "ma_id",
  xlab = "log Odds Ratio",
  transfm = "none"
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  ) +
  scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5))

# Display the plot
print(orchard_plot2)
```

