---
title: "Analysis Code"
author: "Lorenzo Ricolfi"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    toc_depth: 3
    number_sections: no
    theme: cerulean
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

The following packages are required to run the code in this script.
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
#devtools::install_github("EIvimeyCook/ShinyDigitise")
#library(shinyDigitise)
pacman::p_load(tidyr,
               dplyr,
               here,
               ggplot2,
               gridExtra,
               metafor,
               orchaRd,
               stringr
               )       
```

```{r}
ma_details <- read.csv(here("data/ma_details.csv"))
ma_e_details <- read.csv(here("data/ma_e_details.csv"))
```

# Data wrangling and cleaning
```{r}
ma_e_details <- ma_e_details %>%
  filter(chemical_class == "PFAS", 
         !is.na(health_outcome_type),
         ma_es_measure != "unclear",
         !ma_id %in% c("ma_048", "ma_056", "ma_064")) %>%
  mutate(ma_es_measure = recode(
    ma_es_measure,
    "? (regression coefficient) value" = "Beta (regression coefficient) value",
    "beta (regression coefficient) value" = "Beta (regression coefficient) value",
    "Adjusted ? (regression coefficient)" = "Beta (regression coefficient) value",
    "adjusted coefficient regression (?)" = "Beta (regression coefficient) value",
    "relative risk" = "Risk ratio",
    "RR (risk ratio)" = "Risk ratio",
    "OR (odds ratio)" = "Odds ratio",
    "fishers z" = "Fisher's Z",
    "log OR (odds ratio)" = "log Odds ratio")) %>% 
  mutate(chemical_id = recode(
    chemical_id,
    "n-PFOA" = "PFOA",
    "n-PFOS" = "PFOS",
    "PFHsX" = "PFHxS")) %>% 
  left_join(ma_details, by = "ma_id")
```

# Exploratory analyses
## Effect size types
```{r}
es_count <- ma_e_details %>% 
  filter(!is.na(ma_es_measure) & ma_es_measure != "") %>%
  group_by(ma_es_measure) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_chemical_groups = n_distinct(chemical_class),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_health_groups = n_distinct(health_outcome_group),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>% 
  head(20)
# Convert the table to a grid for plotting
es_table <- tableGrob(es_count)
```

```{r, eval=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/es_table.png"), es_table, width = 12, height = 7)
```

## Chemical ID
```{r}
# Count the number of occurrences of each chemical_id
chemical_count <- ma_e_details %>%
  filter(!is.na(chemical_id) & chemical_id != "") %>% 
  group_by(chemical_id) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_health_outcome_group = n_distinct(health_outcome_group),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(20)
# Convert the table to a grid for plotting
chem_table <- tableGrob(chemical_count)
```

```{r, eval=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/chemicals_table.png"), chem_table, width = 11, height = 7)
```

## Health outcomes

```{r}
# Fix encoding issues in health_outcome_type
ma_e_details$health_outcome_type <- iconv(ma_e_details$health_outcome_type, from = "UTF-8", to = "UTF-8", sub = "")

# Count the number of occurrences of each health_outcome_type
outcome_type_table <- ma_e_details %>%
  filter(!is.na(health_outcome_type) & health_outcome_type != "") %>% 
  group_by(health_outcome_type) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate)),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(20)

# Convert the table to a grid for plotting
ho_table <- tableGrob(outcome_type_table)
```


```{r, eval=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/health_outcome_type_table.png"), ho_table, width = 11, height = 6)
```
## Health outcome group
```{r}
# Count the number of occurrences of each health_outcome_info
outcome_group_table <- ma_e_details %>%
  filter(!is.na(health_outcome_group) & health_outcome_group != "") %>% 
  group_by(health_outcome_group) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcome_types = n_distinct(health_outcome_type),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(8)

# Convert the table to a grid for plotting
ho_group_table <- tableGrob(outcome_group_table)
```


```{r, eval=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/health_outcome_group_table.png"), ho_group_table, width = 12, height = 5)
```

# Effect size conversion

See: Borenstein, M., Hedges, L. V., Higgins, J. P., & Rothstein, H. R. (2021). Introduction to meta-analysis. John wiley & sons.

```{r}
# Define a conversion function with an optional baseline risk parameter
convert_to_logOR <- function(effect_measure, point_estimate, ma_l_ci, ma_u_ci, regression_type = NA, baseline_risk = NA) {
  # Ensure the effect measure is treated as character
  effect_measure <- as.character(effect_measure)
  # Ensure point estimate, lower CI, and upper CI are numeric
  point_estimate <- suppressWarnings(as.numeric(as.character(point_estimate)))
  ma_l_ci <- suppressWarnings(as.numeric(as.character(ma_l_ci)))
  ma_u_ci <- suppressWarnings(as.numeric(as.character(ma_u_ci)))
  
  # Conversion for Odds ratio and log Odds ratio
  if (effect_measure == "Odds ratio") {
    # Convert OR to logOR
    return(c(log(point_estimate), log(ma_l_ci), log(ma_u_ci)))
    
  } else if (effect_measure == "log Odds ratio") {
    # Already in logOR
    return(c(point_estimate, ma_l_ci, ma_u_ci))
    
    # Conversion for SMD (standardized mean difference)
  } else if (effect_measure == "SMD (standardized mean difference)") {
    # Following Borenstein et al. (2021): logOR = SMD * pi/sqrt(3)
    logOR_point_estimate <- point_estimate * (pi / sqrt(3))
  
    # Convert lower and upper confidence intervals using the same logic
    logOR_l_ci <- ma_l_ci * (pi / sqrt(3))
    logOR_u_ci <- ma_u_ci * (pi / sqrt(3))
  
    return(c(logOR_point_estimate, logOR_l_ci, logOR_u_ci))

    # Conversion for Relative risk / Risk ratio with optional baseline risk
  } else if (effect_measure == "Risk ratio") {
    if (!is.na(baseline_risk)) {
      # Using the conversion: OR = RR * (1-p0) / (1-RR*p0)
      OR <- point_estimate * (1 - baseline_risk) / (1 - point_estimate * baseline_risk)
      OR_l_ci <- ma_l_ci * (1 - baseline_risk) / (1 - ma_l_ci * baseline_risk)
      OR_u_ci <- ma_u_ci * (1 - baseline_risk) / (1 - ma_u_ci * baseline_risk)
      return(c(log(OR), log(OR_l_ci), log(OR_u_ci)))
    } else {
      # If no baseline risk is provided, fallback to the approximation:
      return(c(log(point_estimate), log(ma_l_ci), log(ma_u_ci)))
    }
    
    # Conversion for Fisher's z (often used for correlations)
  } else if (grepl("fishers z", effect_measure, ignore.case = TRUE)) {
    # Convert z to correlation, then to logOR.
    # r = tanh(z)
    # Approximation: logOR = r * pi/sqrt(3)
    r <- tanh(point_estimate)
    return(r * (pi / sqrt(3)))
    
    # Conversion for regression coefficients assumed from logistic regression:
  } else if (effect_measure %in% c("Beta (regression coefficient) value")) {
    # Check if regression_type is 'logistic'
    if (!is.na(regression_type) && regression_type == "logistic") {
        return(c(point_estimate, ma_l_ci, ma_u_ci))
    } else {
        return(rep(NA, 3)) # Not logistic regression, cannot convert
    }
} else {
    warning("Effect measure conversion not implemented for: ", effect_measure)
    return(rep(NA, 3)) # Return NA for all three
  }
}
```

# Effect size calculation
```{r}
# Apply conversion to each row
ma_e_details <- ma_e_details %>%
  rowwise() %>%
  mutate(
    converted = list(convert_to_logOR(ma_es_measure, ma_point_estimate,
                                     ma_l_ci, ma_u_ci, regression_type, 0.1)),
    logOR = converted[1],
    l_ci_logOR = converted[2],
    u_ci_logOR = converted[3]
  ) %>%
  ungroup() %>%
  select(-converted) %>% 
  filter(!is.na(logOR))
```

# First Research Question
## Overall Model
```{r}
# Calculating effect size standard errors
# SE = (Upper bound – Lower bound) / (2 × 1.96)
ma_e_details <- ma_e_details %>%
  mutate(se = ((u_ci_logOR - l_ci_logOR)/(2*1.96)))  # Convert 95% CI to standard error (se)

# Fitting the multilevel meta-analytic model
mod <- rma.mv(yi = logOR,  # Effect size
              V = se^2, # Variance
              random = list(~1|ma_id, # Meta-analysis identity
                            ~1|ma_e_id), # Meta-analytic point estimate identity
              data = ma_e_details)

# Display model results
summary(mod)
# I²
i2_results <- i2_ml(mod)
print(i2_results)
# Print I² values
cat("I² (between studies):", round(i2_results["I2_ma_id"], 1), "%\n")
cat("I² (within studies):", round(i2_results["I2_ma_e_id"], 1), "%\n")
```

## Forest plot
```{r, eval=FALSE}
ma_e_details$ma_e_id <- factor(ma_e_details$ma_e_id) 
# Reorder ma_e_id levels by logOR (descending), excluding "Overall"
ma_e_details$ma_e_id <- factor(
  ma_e_details$ma_e_id,
  levels = c(
    levels(ma_e_details$ma_e_id)[order(ma_e_details$logOR, decreasing = TRUE)],
    "Overall"  # Ensure "Overall" is last
  )
)

ggplot(ma_e_details, aes(x = logOR, y = ma_e_id)) +
  geom_point(shape = 18, size = 4) +  # Point estimates
  geom_errorbarh(aes(xmin = l_ci_logOR, xmax = u_ci_logOR), 
                 height = 0.2) +      # Confidence intervals
  geom_vline(xintercept = 0,           # Line at null effect
             linetype = "dashed", 
             color = "red") +
  geom_point(data = data.frame(logOR = mod$b[1], 
                              ma_e_id = "Overall",
                              l_ci_logOR = mod$ci.lb,
                              u_ci_logOR = mod$ci.ub),
             shape = 23, size = 4, fill = "red") +
  geom_errorbarh(data = data.frame(logOR = mod$b[1], ma_e_id = "Overall"),
                 aes(xmin = mod$ci.lb, xmax = mod$ci.ub),
                 height = 0.3, size = 1) +
  labs(x = "log Odds Ratio", 
       y = "Meta-analysis",
       title = "Forest Plot of Effect Sizes") +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank()) + # Cleaner horizontal lines
  scale_y_discrete(limits = rev)
```

More elaborated forest plot:
```{r, eval=FALSE}
# Step 1 & 2: Get first point for each ma_id
labels_df <- ma_e_details %>%
    group_by(ma_id) %>%
    slice(1) %>%
    ungroup()

# Step 3: Add geom_text to your plot
ggplot(ma_e_details, aes(x = logOR, y = ma_e_id, color = ma_id)) +
    geom_point(shape = 18, size = 4) +
    geom_errorbarh(aes(xmin = l_ci_logOR, xmax = u_ci_logOR), height = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    geom_point(data = data.frame(logOR = mod$b[1], 
                                 ma_e_id = "Overall",
                                 l_ci_logOR = mod$ci.lb,
                                 u_ci_logOR = mod$ci.ub),
               aes(x = logOR, y = ma_e_id),  # explicitly declare aes
               shape = 23, size = 4, fill = "red", color = "black", inherit.aes = FALSE) +
    geom_errorbarh(data = data.frame(logOR = mod$b[1], ma_e_id = "Overall"),
                   aes(xmin = mod$ci.lb, xmax = mod$ci.ub, y = ma_e_id),
                   height = 0.3, size = 1, inherit.aes = FALSE) +
    # geom_text for labels
    geom_text(data = labels_df,
              aes(x = 1, y = ma_e_id, label = ma_id),  # fixed x value
              hjust = 0,  # left-aligned
              size = 3,
              inherit.aes = FALSE) +
    labs(x = "log Odds Ratio", 
         y = "Meta-analysis",
         title = "Forest Plot of Effect Sizes",
         color = "Meta-analysis ID") +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank()) +
  scale_y_discrete(limits = rev(levels(factor(ma_e_details$ma_e_id))))
```

## OrchaRd
```{r}
# Create orchard plot
orchard_plot <- orchard_plot(
  object = mod,
  mod = "1",
  group = "ma_es_measure",
  xlab = "log Odds Ratio",
  colour  = TRUE,
  transfm = "none"
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  )

# Display the plot
print(orchard_plot)
```


```{r, eval=FALSE}
extreme_logOR <- ma_e_details %>% filter(logOR < -0.5 | logOR > 0.5)
zero_logOR <- ma_e_details %>% filter(logOR == 0)
```

# Impact of effect size metric

```{r}
ma_e_details$ma_es_measure <- as.factor(ma_e_details$ma_es_measure)
ma_e_details$ma_es_measure <- relevel(ma_e_details$ma_es_measure, ref = "Odds ratio")

ma_e_details <- ma_e_details %>%
  mutate(ma_es_measure = recode(
    ma_es_measure,
    "Beta (regression coefficient) value" = "Beta",
    "SMD (standardized mean difference)" = "SMD"))

mod_test <- rma.mv(yi = logOR,  
              V = se^2, 
              random = list(~1|ma_id, 
                            ~1|ma_e_id), 
              mods = ~ ma_es_measure - 1,
              data = ma_e_details)

# Display model results
summary(mod_test)

orchard_plot(
  object = mod_test,
  mod = "ma_es_measure",
  group = "ma_id",
  xlab = "log Odds Ratio",
  colour  = FALSE,
  transfm = "none"
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  )
```

The models showed that our strategy for transformation of effect size metrics into logOR had a significant impact on the overall result.

```{r, eval=FALSE}
ma_e_details_try <- ma_e_details %>% 
  filter(ma_es_measure != "SMD (standardized mean difference)")

mod_try <- rma.mv(yi = logOR,  
              V = se^2, 
              random = list(~1|ma_id, 
                            ~1|ma_e_id), 
              data = ma_e_details_try)

summary(mod_try)

orchard_plot(
  object = mod_try,
  group = "ma_id",
  xlab = "log Odds Ratio",
  colour  = FALSE,
  transfm = "none"
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  )
```

## Excluding transformations

Let's reduce our data set to original effect sizes on the Odds Ratio or log Odds Ratio metrics.

```{r}
ma_e_details_no_transf <- ma_e_details %>% 
  filter(ma_es_measure %in% c("Odds ratio", "log Odds ratio") )

mod_no_transf <- rma.mv(yi = logOR,  
              V = se^2, 
              random = list(~1|ma_id, 
                            ~1|ma_e_id), 
              data = ma_e_details_no_transf)

# Display model results
summary(mod_no_transf)
# I²
i2_results_no_transf <- i2_ml(mod_no_transf)
print(i2_results_no_transf)
# Print I² values
cat("I² (between studies):", round(i2_results_no_transf["I2_ma_id"], 1), "%\n")
cat("I² (within studies):", round(i2_results_no_transf["I2_ma_e_id"], 1), "%\n")

orchard_plot_no_transf <- orchard_plot(
  object = mod_no_transf,
  group = "ma_id",
  xlab = "log Odds Ratio",
  transfm = "none"
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  )

# Display the plot
print(orchard_plot_no_transf)
```

# Second Research Question
## Chemicals

Is there any difference in the significance and strength of the association with adverse health outcomes among individual PFAS?
YES - p-val < .0001

What individual PFAS are significantly associated with adverse health outcomes?
PFBS: 0.2947, p-val = 0.0040
PFHxS: 0.0549, p-val = 0.0040
PFNA: 0.0413, p-val = 0.0742
PFOA: 0.0820, p-val <.0001 
PFOS: 0.0877, p-val <.0001
PFUnDA: -0.0741, 0.0737

```{r}
ma_e_details_no_transf$chemical_id <- as.factor(ma_e_details_no_transf$chemical_id)

mod2 <- rma.mv(yi = logOR,  
              V = se^2, 
              random = list(~1|ma_id, 
                            ~1|ma_e_id), 
              mods = ~ chemical_id -1,
              data = ma_e_details_no_transf)

# Display model results
summary(mod2)
i2_ml(mod2)
```

## Health outcome class

Is there any difference in the significance and strength of the association with PFAS exposure among individual classes (lower level) of adverse health outcomes?
YES - p-val = 0.0005

What classes (lower level) of adverse health outcomes are significantly associated with PFAS exposure?
Hypertensive diseases: p-val = 0.0024
Metabolic disorders: p-val = 0.0055
Oedema, proteinuria, or hypertensive disorders in pregnancy, childbirth, or the puerperium: p-val = 0.0020
Osteopathies or chondropathies: p-val = 0.0410	

```{r}
ma_e_details_no_transf$health_outcome_class <- as.factor(ma_e_details_no_transf$health_outcome_class)
#ma_e_details_no_transf$health_outcome_class <- relevel(ma_e_details_no_transf$health_outcome_class, ref = "Hypertensive diseases")

mod3 <- rma.mv(yi = logOR,  
              V = se^2, 
              random = list(~1|ma_id, 
                            ~1|ma_e_id), 
              mods = ~ health_outcome_class -1,
              data = ma_e_details_no_transf)

# Display model results
summary(mod3)
```

## Healtch outcome group

Is there any difference in the significance and strength of the association with PFAS exposure among individual groups (higher level) of adverse health outcomes?
YES - p-val = 0.0014

What groups (higher level) of adverse health outcome are significantly associated with PFAS exposure?
Diseases of the circulatory system: p-val = 0.0612
Pregnancy, childbirth or the puerperium: p-val = 0.0019

```{r}
ma_e_details_no_transf$health_outcome_group <- as.factor(ma_e_details_no_transf$health_outcome_group)

mod4 <- rma.mv(yi = logOR,  
              V = se^2, 
              random = list(~1|ma_id, 
                            ~1|ma_e_id), 
              mods = ~ health_outcome_group -1,
              data = ma_e_details_no_transf)

# Display model results
summary(mod4)
```

# Third Research Question

## Chemicals


```{r}
ma_e_details_no_transf <- ma_e_details_no_transf %>% 
  filter(chemical_id != "PFAS")

mod_sub <- rma.mv(yi = logOR,  
                  V = se^2,
                  mods = ~ chemical_id -1,
                   random = list(~1|ma_id,
                                 ~1|ma_e_id,
                                 ~ chemical_id|health_outcome_group),
                   test = "t", #to conunteract potential small sample size 
                   sparse = T,
                   data = ma_e_details_no_transf,
                   control = list(iter.max = 1000, 
                                  rel.tol = 1e-8))

# Display model results
summary(mod_sub)
```

```{r}
k_n_data <- ma_e_details_no_transf %>% 
  group_by(chemical_id) %>%
  arrange(chemical_id) %>% 
  summarise(k = n(), n = length(unique(ma_id)))

res_tab1 <- broom::tidy(mod_sub)
res_tab1$ci.lb <- mod_sub$ci.lb
res_tab1$ci.ub <- mod_sub$ci.ub
res_tab1$pval <-mod_sub$pval
res_tab1 <- res_tab1 %>% 
  mutate(
    chemical_id = str_remove(term, "^chemical_id"),  # Remove 'Pchemical_id: ' prefix
  ) %>% 
  arrange(chemical_id)

res_tab1 <- res_tab1 %>% 
  mutate(k = k_n_data$k) %>% 
  mutate(n = k_n_data$n)

# Prepare the data for plotting and order chemical_id by estimate value
plot_data1 <- res_tab1 %>%
  as.data.frame() %>% 
  filter(k > 1) %>% 
  arrange(estimate)

# Create the flipped forest plot
forest_plot1 <- ggplot(plot_data1, aes(y = reorder(chemical_id, estimate), x = estimate)) +
  geom_point(aes(size = k),  # Make point size proportional to k
             color = ifelse(plot_data1$pval > 0.05, "grey", "black"),  # You can change the color if desired
             alpha = 1) +  # Add transparency for better visibility 
  scale_size(range = c(5, 13)) + # Set minimum size to 3 and maximum to 6
  geom_errorbarh(aes(xmin = ci.lb, xmax = ci.ub), 
                 height = 0.2,
                 alpha = ifelse(plot_data1$pval > 0.05, 0.6, 1)) +  # Horizontal error bars
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +  # Reference line at 1
  geom_text(aes(label = paste0("k = ", k, " (", n, ")"), x = - 0.3),
            hjust = 0, size = 5, # Add k values as text next to the points
            fontface = ifelse(plot_data1$pval < 0.05, "bold", "plain")) +  # Access pval outside aes()
  # Add "*" next to PFUnDA and PFNA
  # geom_text(data = subset(plot_data1, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = -0.21, label = "*"), size = 9, color = "black") +
  # geom_text(data = subset(plot_data1, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = ci.lb, label = "*"),
  #           hjust = 1.2, vjust = 0.7, size = 6, color = "black") +
  xlab("logOR") +
  ylab("PFAS") +
  #ggtitle("B") +
  theme_minimal() +  # Use a minimal theme
  theme(legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
         plot.title = element_text(face = "bold", size = 18),
       legend.title = element_text(size = 16),
        legend.text = element_text(size = 15),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.text.y = element_text(size = 17),
       axis.text.x = element_text(size = 17)) #+  
   #scale_x_break(c(4, 6)) +
   #scale_x_continuous(breaks = seq(0, 4, by = 1), limits = c(-0.21, 4))
  
forest_plot1
```

## Health outcomes (groups)

```{r}
mod_sub2 <- rma.mv(yi = logOR,  
                  V = se^2,
                  mods = ~ health_outcome_group -1,
                   random = list(~1|ma_id,
                                 ~1|ma_e_id,
                                 ~ health_outcome_group|chemical_id),
                   test = "t", #to conunteract potential small sample size 
                   sparse = T,
                   data = ma_e_details_no_transf,
                   control = list(iter.max = 1000, 
                                  rel.tol = 1e-8))

# Display model results
summary(mod_sub2)
```

```{r}
k_n_data <- ma_e_details_no_transf %>% 
  group_by(health_outcome_group) %>%
  arrange(health_outcome_group) %>% 
  summarise(k = n(), n = length(unique(ma_id)))

res_tab1 <- broom::tidy(mod_sub2)
res_tab1$ci.lb <- mod_sub2$ci.lb
res_tab1$ci.ub <- mod_sub2$ci.ub
res_tab1$pval <-mod_sub2$pval
res_tab1 <- res_tab1 %>% 
  mutate(
    health_outcome_group = str_remove(term, "^health_outcome_group"),  # Remove 'Pchemical_id: ' prefix
  ) %>% 
  arrange(health_outcome_group)

res_tab1 <- res_tab1 %>% 
  mutate(k = k_n_data$k) %>% 
  mutate(n = k_n_data$n)

# Prepare the data for plotting and order health_outcome_group by estimate value
plot_data1 <- res_tab1 %>%
  as.data.frame() %>% 
  #filter(k > 1) %>% 
  filter(n > 1) %>% 
  arrange(estimate)

# Create the flipped forest plot
forest_plot2 <- ggplot(plot_data1, aes(y = reorder(health_outcome_group, estimate), x = estimate)) +
  geom_point(aes(size = k),  # Make point size proportional to k
             color = ifelse(plot_data1$pval > 0.05, "grey", "black"),  # You can change the color if desired
             alpha = 1) +  # Add transparency for better visibility 
  scale_size(range = c(5, 13)) + # Set minimum size to 3 and maximum to 6
  geom_errorbarh(aes(xmin = ci.lb, xmax = ci.ub), 
                 height = 0.2,
                 alpha = ifelse(plot_data1$pval > 0.05, 0.6, 1)) +  # Horizontal error bars
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +  # Reference line at 1
  geom_text(aes(label = paste0("k = ", k, " (", n, ")"), x = - 0.3),
            hjust = 0, size = 5, # Add k values as text next to the points
            fontface = ifelse(plot_data1$pval < 0.05, "bold", "plain")) +  # Access pval outside aes()
  # Add "*" next to PFUnDA and PFNA
  # geom_text(data = subset(plot_data1, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = -0.21, label = "*"), size = 9, color = "black") +
  # geom_text(data = subset(plot_data1, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = ci.lb, label = "*"),
  #           hjust = 1.2, vjust = 0.7, size = 6, color = "black") +
  xlab("logOR") +
  ylab("Health Outcome Group") +
  ggtitle("A") +
  theme_minimal() +  # Use a minimal theme
  theme(legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
         plot.title = element_text(face = "bold", size = 18),
       legend.title = element_text(size = 16),
        legend.text = element_text(size = 15),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.text.y = element_text(size = 17),
       axis.text.x = element_text(size = 17)) #+  
   #scale_x_break(c(4, 6)) +
   #scale_x_continuous(breaks = seq(0, 4, by = 1), limits = c(-0.21, 4))
  
forest_plot2
```

## Health outcomes (classes)

```{r}
mod_sub3 <- rma.mv(yi = logOR,  
                  V = se^2,
                  mods = ~ health_outcome_class -1,
                   random = list(~1|ma_id,
                                 ~1|ma_e_id,
                                 ~ health_outcome_class|chemical_id),
                   test = "t", #to conunteract potential small sample size 
                   sparse = T,
                   data = ma_e_details_no_transf,
                   control = list(iter.max = 1000, 
                                  rel.tol = 1e-8))

# Display model results
summary(mod_sub3)
```

```{r}
k_n_data <- ma_e_details_no_transf %>% 
  group_by(health_outcome_class) %>%
  arrange(health_outcome_class) %>% 
  summarise(k = n(), n = length(unique(ma_id)))

res_tab1 <- broom::tidy(mod_sub3)
res_tab1$ci.lb <- mod_sub3$ci.lb
res_tab1$ci.ub <- mod_sub3$ci.ub
res_tab1$pval <-mod_sub3$pval
res_tab1 <- res_tab1 %>% 
  mutate(
    health_outcome_class = str_remove(term, "^health_outcome_class"),  # Remove 'Pchemical_id: ' prefix
  ) %>% 
  arrange(health_outcome_class)

res_tab1 <- res_tab1 %>% 
  mutate(k = k_n_data$k) %>% 
  mutate(n = k_n_data$n)

# Prepare the data for plotting and order health_outcome_class by estimate value
plot_data1 <- res_tab1 %>%
  as.data.frame() %>% 
  #filter(k > 1) %>% 
  filter(n > 1) %>% 
  arrange(estimate)

# Create the flipped forest plot
forest_plot2 <- ggplot(plot_data1, aes(y = reorder(health_outcome_class, estimate), x = estimate)) +
  geom_point(aes(size = k),  # Make point size proportional to k
             color = ifelse(plot_data1$pval > 0.05, "grey", "black"),  # You can change the color if desired
             alpha = 1) +  # Add transparency for better visibility 
  scale_size(range = c(5, 13)) + # Set minimum size to 3 and maximum to 6
  geom_errorbarh(aes(xmin = ci.lb, xmax = ci.ub), 
                 height = 0.2,
                 alpha = ifelse(plot_data1$pval > 0.05, 0.6, 1)) +  # Horizontal error bars
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +  # Reference line at 1
  geom_text(aes(label = paste0("k = ", k, " (", n, ")"), x = - 0.3),
            hjust = 0, size = 5, # Add k values as text next to the points
            fontface = ifelse(plot_data1$pval < 0.05, "bold", "plain")) +  # Access pval outside aes()
  # Add "*" next to PFUnDA and PFNA
  # geom_text(data = subset(plot_data1, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = -0.21, label = "*"), size = 9, color = "black") +
  # geom_text(data = subset(plot_data1, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = ci.lb, label = "*"),
  #           hjust = 1.2, vjust = 0.7, size = 6, color = "black") +
  xlab("logOR") +
  ylab("Health Outcome Class") +
  ggtitle("B") +
  theme_minimal() +  # Use a minimal theme
  theme(legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
         plot.title = element_text(face = "bold", size = 18),
       legend.title = element_text(size = 16),
        legend.text = element_text(size = 15),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.text.y = element_text(size = 17),
       axis.text.x = element_text(size = 17)) #+  
   #scale_x_break(c(4, 6)) +
   #scale_x_continuous(breaks = seq(0, 4, by = 1), limits = c(-0.21, 4))
  
forest_plot2
```