---
title: "Analysis Code"
author: "Lorenzo Ricolfi"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    toc_depth: 3
    number_sections: no
    theme: cerulean
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following analysis steps were removed from the present file because out of its scope:

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE, include=FALSE}
#devtools::install_github("EIvimeyCook/ShinyDigitise")
#library(shinyDigitise)
pacman::p_load(tidyr,
               dplyr,
               here,
               ggplot2,
               gridExtra,
               metafor,
               orchaRd,
               stringr
               )       
```

- Data wrangling and cleaning
```{r, include=FALSE}
ma_details <- read.csv(here("data/ma_details.csv"))
ma_e_details <- read.csv(here("data/ma_e_details.csv"))

ma_e_details <- ma_e_details %>%
  filter(chemical_class == "PFAS", 
         !is.na(health_outcome_type),
         ma_es_measure != "unclear",
         !ma_id %in% c("ma_048", "ma_056", "ma_064")) %>%
  mutate(ma_es_measure = recode(
    ma_es_measure,
    "? (regression coefficient) value" = "Beta (regression coefficient) value",
    "beta (regression coefficient) value" = "Beta (regression coefficient) value",
    "Adjusted ? (regression coefficient)" = "Beta (regression coefficient) value",
    "adjusted coefficient regression (?)" = "Beta (regression coefficient) value",
    "relative risk" = "Risk ratio",
    "RR (risk ratio)" = "Risk ratio",
    "OR (odds ratio)" = "Odds ratio",
    "fishers z" = "Fisher's Z",
    "log OR (odds ratio)" = "log Odds ratio")) %>% 
  mutate(chemical_id = recode(
    chemical_id,
    "n-PFOA" = "PFOA",
    "n-PFOS" = "PFOS",
    "PFHsX" = "PFHxS")) %>% 
  left_join(ma_details, by = "ma_id")
```

- Exploratory analyses

```{r, include=FALSE}
es_count <- ma_e_details %>% 
  filter(!is.na(ma_es_measure) & ma_es_measure != "") %>%
  group_by(ma_es_measure) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_chemical_groups = n_distinct(chemical_class),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_health_groups = n_distinct(health_outcome_group),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>% 
  head(20)
# Convert the table to a grid for plotting
es_table <- tableGrob(es_count)
```

```{r, eval=FALSE, include=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/es_table.png"), es_table, width = 12, height = 7)
```


```{r, include=FALSE}
# Count the number of occurrences of each chemical_id
chemical_count <- ma_e_details %>%
  filter(!is.na(chemical_id) & chemical_id != "") %>% 
  group_by(chemical_id) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_health_outcome_group = n_distinct(health_outcome_group),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(20)
# Convert the table to a grid for plotting
chem_table <- tableGrob(chemical_count)
```

```{r, eval=FALSE, include=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/chemicals_table.png"), chem_table, width = 11, height = 7)
```

```{r, include=FALSE}
# Fix encoding issues in health_outcome_type
ma_e_details$health_outcome_type <- iconv(ma_e_details$health_outcome_type, from = "UTF-8", to = "UTF-8", sub = "")

# Count the number of occurrences of each health_outcome_type
outcome_type_table <- ma_e_details %>%
  filter(!is.na(health_outcome_type) & health_outcome_type != "") %>% 
  group_by(health_outcome_type) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcomes = n_distinct(health_outcome_type),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate)),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(20)

# Convert the table to a grid for plotting
ho_table <- tableGrob(outcome_type_table)
```

```{r, eval=FALSE, include=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/health_outcome_type_table.png"), ho_table, width = 11, height = 6)
```

```{r, include=FALSE}
# Count the number of occurrences of each health_outcome_info
outcome_group_table <- ma_e_details %>%
  filter(!is.na(health_outcome_group) & health_outcome_group != "") %>% 
  group_by(health_outcome_group) %>%
  summarise(n_meta_analyses = n_distinct(ma_id),
            n_health_outcome_types = n_distinct(health_outcome_type),
            n_non_na_point_estimates = sum(!is.na(ma_point_estimate))) %>%
  arrange(desc(n_meta_analyses)) %>%
  head(8)

# Convert the table to a grid for plotting
ho_group_table <- tableGrob(outcome_group_table)
```

```{r, eval=FALSE, include=FALSE}
# Save the table as an image
ggsave(here("figs/exploratory_analysis/health_outcome_group_table.png"), ho_group_table, width = 12, height = 5)
```

- Effect size conversion & calculation

```{r, include=FALSE}
# conversion function to transform effect size metrics into logOR
source(here("function", "conversion.R"))
```

```{r, include=FALSE}
# Apply conversion to each row
ma_e_details <- ma_e_details %>%
  rowwise() %>%
  mutate(
    converted = list(convert_to_logOR(ma_es_measure, ma_point_estimate,
                                     ma_l_ci, ma_u_ci, regression_type, 0.1)),
    logOR = converted[1],
    l_ci_logOR = converted[2],
    u_ci_logOR = converted[3]
  ) %>%
  ungroup() %>%
  select(-converted) %>% 
  filter(!is.na(logOR))
```

# First Research Question
**Question:** How does exposure to per- and polyfluoroalkyl substances (PFAS) affect human health?

**Objective:** To evaluate the magnitude and variability of associations between PFAS exposure and adverse human health outcomes based on multiple meta-analyses.

BUT, 

Before looking at the meta-analytic model, I have to tell you that:

- I run a meta-analytic model including all effect sizes and then tested their impact on the overall result. 

- I found that the strategy we used to convert effect size metrics into logOR has an impact on the overall results.

- Thus, I had to trim the dataset to only original logOR and OR effect size metrics.

```{r, warning=FALSE, message=FALSE, include=FALSE}
# Calculating effect size standard errors
# SE = (Upper bound – Lower bound) / (2 × 1.96)
ma_e_details <- ma_e_details %>%
  mutate(se = ((u_ci_logOR - l_ci_logOR)/(2*1.96)))  # Convert 95% CI to standard error (se)

# Fitting the multilevel meta-analytic model
mod <- rma.mv(yi = logOR,  # Effect size
              V = se^2, # Variance
              random = list(~1|ma_id, # Meta-analysis identity
                            ~1|ma_e_id), # Meta-analytic point estimate identity
              data = ma_e_details)

# Display model results
summary(mod)
# I²
i2_results <- i2_ml(mod)
print(i2_results)
# Print I² values
cat("I² (between studies):", round(i2_results["I2_ma_id"], 1), "%\n")
cat("I² (within studies):", round(i2_results["I2_ma_e_id"], 1), "%\n")
```

```{r, include=FALSE, warning=FALSE}
# Create orchard plot
orchard_plot <- orchard_plot(
  object = mod,
  mod = "1",
  group = "ma_es_measure",
  xlab = "log Odds Ratio",
  colour  = TRUE,
  transfm = "none"
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  )

# Display the plot
print(orchard_plot)
```

```{r, eval=FALSE, include=FALSE}
extreme_logOR <- ma_e_details %>% filter(logOR < -0.5 | logOR > 0.5)
zero_logOR <- ma_e_details %>% filter(logOR == 0)
```

Here we can see the impact of effect size metrics (moderator) and the reason why I decided to trim the dataset to only original logOR and OR effect size metrics:

```{r, include=FALSE, warning=FALSE}
ma_e_details$ma_es_measure <- as.factor(ma_e_details$ma_es_measure)
ma_e_details$ma_es_measure <- relevel(ma_e_details$ma_es_measure, ref = "Odds ratio")

ma_e_details <- ma_e_details %>%
  mutate(ma_es_measure = recode(
    ma_es_measure,
    "Beta (regression coefficient) value" = "Beta",
    "SMD (standardized mean difference)" = "SMD"))
```

```{r, results=FALSE, warning=FALSE}
mod_es <- rma.mv(yi = logOR,  
              V = se^2, 
              random = list(~1|ma_id, 
                            ~1|ma_e_id), 
              mods = ~ ma_es_measure - 1,
              data = ma_e_details)
```

```{r, include=FALSE, warning=FALSE}
# Display model results
summary(mod_es)
```

```{r, echo=FALSE, warning=FALSE}
orchard_plot(
  object = mod_es,
  mod = "ma_es_measure",
  group = "ma_id",
  xlab = "log Odds Ratio",
  colour  = FALSE,
  transfm = "none"
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  )
```

## The Meta-analytic Model

We are fitting a multilevel random-effects model via *rma.mv()* from the *metafor* package.

Random effect structure:

- **~1 | ma_id** introduces a random intercept for each meta-analysis. It acknowledges that different meta-analyses may have different baseline effects due to design, populations, etc.

- **~1 | ma_e_id** introduces a random intercept for each effect size estimate within a meta-analysis. It accounts for non-independence of estimates within the same meta-analysis.

```{r, results=FALSE}
dat <- ma_e_details %>% 
  filter(ma_es_measure %in% c("Odds ratio", "log Odds ratio") )

mod_no_transf <- rma.mv(yi = logOR,  
                        V = se^2, 
                        random = list(~1|ma_id, 
                                      ~1|ma_e_id), 
                        data = dat,
                        test = "t")

summary(mod_no_transf)
```

```{r,echo=FALSE}
orchard_plot_no_transf <- orchard_plot(
  object = mod_no_transf,
  group = "ma_id",
  xlab = "log Odds Ratio",
  transfm = "none",
  trunk.size = 0.8,
  branch.size = 2.5,
  alpha = 0.8
) + 
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.3),
    axis.text.x = element_text(size = 10)
  )

# Display the plot
print(orchard_plot_no_transf)

```

## Analysis of Heterogeneity

```{r, echo=FALSE}
# I²
i2_results_no_transf <- i2_ml(mod_no_transf)
print(i2_results_no_transf)
# Print I² values
cat("I² (between studies):", round(i2_results_no_transf["I2_ma_id"], 1), "%\n")
cat("I² (within studies):", round(i2_results_no_transf["I2_ma_e_id"], 1), "%\n")
```

```{r, include=FALSE}
# Back-transforming logOR to the Odds Ratio:
# estimate
exp(0.0600)
# lower 95% confidence interval
exp(0.0345)
# higher 95% confidence interval
exp(0.0855)
```

## Summary Results

**1)**  Exposure to PFAS showed a weak but significant association with increased odds of an adverse health outcome, across all meta-analyses included. Specifically, a 6% (CI 3.5% - 8.9%) increase.

**2)**  Most of the variability in logORs comes from within meta-analyses (66.6%).


# Second Research Question
## Chemicals

**Question:** Is there any difference in the significance and strength of the association with adverse health outcomes among individual PFAS?

**Question:** What individual PFAS are significantly associated with adverse health outcomes?

```{r, include=FALSE}
dat$chemical_id <- as.factor(dat$chemical_id)
```

```{r, results=FALSE}
mod2 <- rma.mv(yi = logOR,  
               V = se^2,
               mods = ~ chemical_id -1,
               random = list(~1|ma_id,
                             ~1|ma_e_id),
               test = "t", 
               sparse = T,
               data = dat,
               control = list(iter.max = 1000, 
                              rel.tol = 1e-8))

summary(mod2)
```

```{r, echo=FALSE, out.height=600, out.width=1000, fig.width=11, fig.height=8}
k_n_data <- dat %>% 
  group_by(chemical_id) %>%
  arrange(chemical_id) %>% 
  summarise(k = n(), n = length(unique(ma_id)))

res_tab1 <- broom::tidy(mod2)
res_tab1$ci.lb <- mod2$ci.lb
res_tab1$ci.ub <- mod2$ci.ub
res_tab1$pval <-mod2$pval
res_tab1 <- res_tab1 %>% 
  mutate(
    chemical_id = str_remove(term, "^chemical_id"),  # Remove 'Pchemical_id: ' prefix
  ) %>% 
  arrange(chemical_id)

res_tab1 <- res_tab1 %>% 
  mutate(k = k_n_data$k) %>% 
  mutate(n = k_n_data$n)

# Prepare the data for plotting and order chemical_id by estimate value
plot_data1 <- res_tab1 %>%
  as.data.frame() %>% 
  filter(n > 1) %>% 
  arrange(estimate)

# Create the flipped forest plot
forest_plot1 <- ggplot(plot_data1, aes(y = reorder(chemical_id, estimate), x = estimate)) +
  geom_point(aes(size = k),  # Make point size proportional to k
             color = ifelse(plot_data1$pval > 0.05, "grey", "black"), 
             alpha = 1) +  # Add transparency for better visibility 
  scale_size(range = c(5, 13)) + # Set minimum size to 3 and maximum to 6
  geom_errorbarh(aes(xmin = ci.lb, xmax = ci.ub), 
                 height = 0.2,
                 alpha = ifelse(plot_data1$pval > 0.05, 0.6, 1)) +  # Horizontal error bars
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +  # Reference line at 1
  geom_text(aes(label = paste0("k = ", k, " (", n, ")"), x = - 0.3),
            hjust = 0, size = 5, # Add k values as text next to the points
            fontface = ifelse(plot_data1$pval < 0.05, "bold", "plain")) +  # Access pval outside aes()
  # Add "*" next to PFUnDA and PFNA
  # geom_text(data = subset(plot_data1, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = -0.21, label = "*"), size = 9, color = "black") +
  # geom_text(data = subset(plot_data1, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = ci.lb, label = "*"),
  #           hjust = 1.2, vjust = 0.7, size = 6, color = "black") +
  xlab("logOR") +
  ylab("PFAS") +
  #ggtitle("B") +
  theme_minimal() +  # Use a minimal theme
  theme(legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
         plot.title = element_text(face = "bold", size = 16),
       legend.title = element_text(size = 14),
        legend.text = element_text(size = 13),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.y = element_text(size = 15),
       axis.text.x = element_text(size = 15)) #+  
   #scale_x_break(c(4, 6)) +
   #scale_x_continuous(breaks = seq(0, 4, by = 1), limits = c(-0.21, 4))
  
forest_plot1
```

## Health Outcome Class

**Question:** Is there any difference in the significance and strength of the association with PFAS exposure among individual classes (lower level) of adverse health outcomes?

**Question:** What classes (lower level) of adverse health outcomes are significantly associated with PFAS exposure?

```{r, include=FALSE}
dat$health_outcome_class <- as.factor(dat$health_outcome_class)
#dat$health_outcome_class <- relevel(dat$health_outcome_class, ref = "Hypertensive diseases")

dat <- dat %>% 
mutate(health_outcome_class = recode(
    health_outcome_class,
    "Oedema, proteinuria, or hypertensive disorders in pregnancy, childbirth, or the puerperium" = "Hypertensive disorders \n in pregnancy",
    "Abortive outcome of pregnancy" = "Abortive outcome \n of pregnancy",
    "Certain specified maternal disorders predominantly related to pregnancy" = "Maternal disorders \n predominantly related \n to pregnancy",
    "Disorders of newborn related to length of gestation or fetal growth" = "Disorders of newborn \n related to length of \n gestation or fetal growth",
    "Diseases of the female genital system" = "Diseases of the \n female genital system"
  ))
```


```{r, results=FALSE}
mod3 <- rma.mv(yi = logOR,  
                  V = se^2,
                  mods = ~ health_outcome_class -1,
                  random = list(~1|ma_id,
                                ~1|ma_e_id),
                  test = "t", 
                  sparse = T,
                  data = dat,
                  control = list(iter.max = 1000, 
                                  rel.tol = 1e-8))

summary(mod3)
```

```{r,echo=FALSE, out.height=600, out.width=1000, fig.width=11, fig.height=8}
k_n_data2 <- dat %>% 
  group_by(health_outcome_class) %>%
  arrange(health_outcome_class) %>% 
  summarise(k = n(), n = length(unique(ma_id)))

res_tab2 <- broom::tidy(mod3)
res_tab2$ci.lb <- mod3$ci.lb
res_tab2$ci.ub <- mod3$ci.ub
res_tab2$pval <-mod3$pval
res_tab2 <- res_tab2 %>% 
  mutate(
    health_outcome_class = str_remove(term, "^health_outcome_class"),  # Remove 'Pchemical_id: ' prefix
  ) %>% 
  arrange(health_outcome_class)

res_tab2 <- res_tab2 %>% 
  mutate(k = k_n_data2$k) %>% 
  mutate(n = k_n_data2$n)

# Prepare the data for plotting and order health_outcome_class by estimate value
plot_data2 <- res_tab2 %>%
  as.data.frame() %>% 
  #filter(k > 1) %>% 
  filter(n > 1) %>% 
  arrange(estimate)

# Create the flipped forest plot
forest_plot2 <- ggplot(plot_data2, aes(y = reorder(health_outcome_class, estimate), x = estimate)) +
  geom_point(aes(size = k),  # Make point size proportional to k
             color = ifelse(plot_data2$pval > 0.05, "grey", "black"), 
             alpha = 1) +  # Add transparency for better visibility 
  scale_size(range = c(5, 13)) + # Set minimum size to 3 and maximum to 6
  geom_errorbarh(aes(xmin = ci.lb, xmax = ci.ub), 
                 height = 0.2,
                 alpha = ifelse(plot_data2$pval > 0.05, 0.6, 1)) +  # Horizontal error bars
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +  # Reference line at 1
  geom_text(aes(label = paste0("k = ", k, " (", n, ")"), x = - 0.3),
            hjust = 0, size = 5, # Add k values as text next to the points
            fontface = ifelse(plot_data2$pval < 0.05, "bold", "plain")) +  # Access pval outside aes()
  # Add "*" next to PFUnDA and PFNA
  # geom_text(data = subset(plot_data2, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = -0.21, label = "*"), size = 9, color = "black") +
  # geom_text(data = subset(plot_data2, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = ci.lb, label = "*"),
  #           hjust = 1.2, vjust = 0.7, size = 6, color = "black") +
  xlab("logOR") +
  ylab("Health Outcome Class") +
  ggtitle("A") +
  theme_minimal() +  # Use a minimal theme
  theme(legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
         plot.title = element_text(face = "bold", size = 18),
       legend.title = element_text(size = 16),
        legend.text = element_text(size = 15),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.text.y = element_text(size = 17),
       axis.text.x = element_text(size = 17)) #+  
   #scale_x_break(c(4, 6)) +
   #scale_x_continuous(breaks = seq(0, 4, by = 1), limits = c(-0.21, 4))
  
forest_plot2
```

## Health Outcome Group

**Question:** Is there any difference in the significance and strength of the association with PFAS exposure among individual groups (higher level) of adverse health outcomes?

**Question:** What groups (higher level) of adverse health outcome are significantly associated with PFAS exposure?

```{r, include=FALSE}
dat$health_outcome_group <- as.factor(dat$health_outcome_group)

dat <- dat %>% 
mutate(health_outcome_group = recode(
    health_outcome_group,
    "Endocrine, nutritional or metabolic diseases" = "Endocrine diseases",
    "Diseases of the circulatory system" = "Diseases of the\ncirculatory system",
    "Pregnancy, childbirth or the puerperium" = "Pregnancy or\nchildbirth",
    "Certain conditions originating in the perinatal period" = "Perinatal period",
    "Mental, behavioural or neurodevelopmental disorders" = "Neurodevelopmental\ndisorders",
    "Diseases of the musculoskeletal system or connective tissue" = "Diseases of the\nmusculoskeletal system",
    "Symptoms, signs or clinical findings, not elsewhere classified" = "Symptoms not\nelsewhere classified",
    "Diseases of the respiratory system" = "Diseases of the\nrespiratory system",
    "Diseases of the genitourinary system" = "Diseases of the\ngenitourinary system"
  ))
```


```{r, results=FALSE}
mod4 <- rma.mv(yi = logOR,  
                  V = se^2,
                  mods = ~ health_outcome_group -1,
                   random = list(~1|ma_id,
                                 ~1|ma_e_id),
                   test = "t",
                   sparse = T,
                   data = dat,
                   control = list(iter.max = 1000, 
                                  rel.tol = 1e-8))

summary(mod4)
```

```{r, echo=FALSE, out.height=600, out.width=1000, fig.width=11, fig.height=8}
k_n_data3 <- dat %>% 
  group_by(health_outcome_group) %>%
  arrange(health_outcome_group) %>% 
  summarise(k = n(), n = length(unique(ma_id)))

res_tab3 <- broom::tidy(mod4)
res_tab3$ci.lb <- mod4$ci.lb
res_tab3$ci.ub <- mod4$ci.ub
res_tab3$pval <- mod4$pval
res_tab3 <- res_tab3 %>% 
  mutate(
    health_outcome_group = str_remove(term, "^health_outcome_group"),  # Remove 'Pchemical_id: ' prefix
  ) %>% 
  arrange(health_outcome_group)

res_tab3 <- res_tab3 %>% 
  mutate(k = k_n_data3$k) %>% 
  mutate(n = k_n_data3$n)

# Prepare the data for plotting and order health_outcome_group by estimate value
plot_data3 <- res_tab3 %>%
  as.data.frame() %>% 
  #filter(k > 1) %>% 
  filter(n > 1) %>% 
  arrange(estimate)

# Create the flipped forest plot
forest_plot3 <- ggplot(plot_data3, aes(y = reorder(health_outcome_group, estimate), x = estimate)) +
  geom_point(aes(size = k),  # Make point size proportional to k
             color = ifelse(plot_data3$pval > 0.05, "grey", "black"),
             alpha = 1) +  # Add transparency for better visibility 
  scale_size(range = c(5, 13)) + # Set minimum size to 3 and maximum to 6
  geom_errorbarh(aes(xmin = ci.lb, xmax = ci.ub), 
                 height = 0.2,
                 alpha = ifelse(plot_data3$pval > 0.05, 0.6, 1)) +  # Horizontal error bars
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +  # Reference line at 1
  geom_text(aes(label = paste0("k = ", k, " (", n, ")"), x = - 0.3),
            hjust = 0, size = 5, # Add k values as text next to the points
            fontface = ifelse(plot_data3$pval < 0.05, "bold", "plain")) +  # Access pval outside aes()
  # Add "*" next to PFUnDA and PFNA
  # geom_text(data = subset(plot_data3, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = -0.21, label = "*"), size = 9, color = "black") +
  # geom_text(data = subset(plot_data3, PFAS_ID %in% c("PFUnDA", "PFNA", "PFTrDA", "PFDoDA", "FOSA", "PFHxS", "PFDS", "PFTeDA", "PFOA", "PFBA", "PFPeA", "PFHpA", "PFBS", "PFHxA")),
  #           aes(x = ci.lb, label = "*"),
  #           hjust = 1.2, vjust = 0.7, size = 6, color = "black") +
  xlab("logOR") +
  ylab("Health Outcome Group") +
  ggtitle("B") +
  theme_minimal() +  # Use a minimal theme
  theme(legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
         plot.title = element_text(face = "bold", size = 18),
       legend.title = element_text(size = 16),
        legend.text = element_text(size = 15),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.text.y = element_text(size = 17),
       axis.text.x = element_text(size = 17)) #+  
   #scale_x_break(c(4, 6)) +
   #scale_x_continuous(breaks = seq(0, 4, by = 1), limits = c(-0.21, 4))
  
forest_plot3
```


## Summary Results

**1)**  We found statistically significant differences in the associations between individual PFAS and adverse health outcomes (p < 0.0001).

**2)**  The following substances showed significant associations with adverse health outcomes: 

-  PFBS: 0.2947, p-val = 0.0040
-  PFHxS: 0.0549, p-val = 0.0040 
-  PFNA: 0.0413, p-val = 0.0742 
-  PFOA: 0.0820, p-val <.0001
-  PFOS: 0.0877, p-val <.0001 
-  PFUnDA: -0.0741, p-val = 0.0737

**3)**  We also found significant differences in the associations between PFAS exposure and specific classes of adverse health outcomes (p = 0.0005). Classes is the lower level of classification according to the International Statistical Classification of Diseases (ICD) 2024 version – 11th revision.

**4)**  The following outcome classes were significantly associated with PFAS exposure:
-  Hypertensive diseases: p-val = 0.0024 
-  Metabolic disorders: p-val = 0.0055 
-  Oedema, proteinuria, or hypertensive disorders in pregnancy, childbirth, or the puerperium: p-val = 0.0020 
-  Osteopathies or chondropathies: p-val = 0.0410

**5)**  Statistically significant differences were also observed in the associations between PFAS exposure and broader groups of adverse health outcomes (p = 0.0014). Groups is the higher level of classification according to the International Statistical Classification of Diseases (ICD) 2024 version – 11th revision.

**6)**  The following health outcome groups showed significant associations with PFAS exposure:

-  Diseases of the circulatory system: p-val = 0.0612 
-  Pregnancy, childbirth or the puerperium: p-val = 0.0019


# Third Research Question

**Question:** What are the significant associations between individual PFAS and specific health outcomes, and how strong are these correlations?

**Objective:** To determine significant PFAS-health outcome correlations and assess the strength of these relationships.

BUT

Before addressing our third research question through quantitative analysis, we first perform a qualitative assessment to determine the number of meta-analyses and effect sizes available for each combination of individual PFAS and health outcome group.

## Qualitative Synthesis

```{r,echo=FALSE, out.height=600, out.width=1000, fig.width=11, fig.height=10}
# custom function
source(here("function", "custom.R"))

# get estimate for each cell
est_dat <- dat %>% group_by(chemical_id, health_outcome_group) %>%
  group_modify(~ custom_meta_aggregate(.x, rho = 0.5)) %>% ungroup()

# traditional map with the number of study
est_dat$chemical_id <- as.factor(est_dat$chemical_id)

Box1_map1 <- ggplot(est_dat, aes(x = chemical_id, y = health_outcome_group, size = n_studies) ) +
  geom_point(alpha = 0.5, color = RColorBrewer::brewer.pal(n = 8, name = "Dark2")[1]) + 
  labs(x = "PFAS", y = "Health Outcome Group") +
  scale_size(range=c(5,10)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + # 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  guides(size = "none") +
  theme(legend.position='top', 
        legend.justification='right',
        legend.direction='horizontal', 
        axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black")) +
  geom_text(aes(label = as.character(n_studies)), size = 4, color = "gray10") +
  labs(caption = "The value in the cell is the number of studies") + 
   theme(plot.caption = element_text(size = 10, color = "gray10", face = "italic"),
         axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text = element_text(size = 11),
         axis.title = element_text(size = 12, face = "bold")) 

Box1_map1
```

```{r,echo=FALSE, out.height=600, out.width=1000, fig.width=11, fig.height=10}
# traditional map with the number of effect size
Box1_map2 <- ggplot(est_dat, aes(x = chemical_id, y = health_outcome_group, size = n_es)) +
  geom_point(alpha = 0.5, color = RColorBrewer::brewer.pal(n = 8, name = "Dark2")[3]) + 
  labs(x = "PFAS", y = "Health Outcome Group") +
  scale_size(range=c(5,10)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  guides(size = "none") +
  theme(legend.position='top', 
        legend.justification='right',
        legend.direction='horizontal',  
        axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black")) +
  geom_text(aes(label = as.character(n_es)), size = 4, color = "gray10") +
  labs(caption = "The value in the cell is the number of effect sizes") +
   theme(plot.caption = element_text(size = 10, color = "gray10", face = "italic"),
         axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text = element_text(size = 12),
         axis.title = element_text(face = "bold", size = 12))

Box1_map2
```

```{r, include=FALSE, out.height=600, out.width=1000, fig.width=11, fig.height=10}
# with effect size information
Box1_map3 <- ggplot(est_dat, aes(x = chemical_id, y = health_outcome_group, size = n_es, color = estimate)) +
  geom_point(alpha = 0.6) + 
  scale_color_gradient(
    low = "blue",
    high = "red", 
    limits = c(-1.5,1),
    guide = "colourbar") + 
  labs(x = "PFAS", y = "Health Outcome Group", color = "Meta-analytic mean effect size") +
  scale_size(range=c(5,10)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  guides(size = "none") +
  theme(legend.position = 'top', 
        legend.justification = 'left',
        legend.direction = 'horizontal', 
        axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black")) +
  geom_text(aes(label = as.character(n_es)), size = 4, color = "gray10") +
  labs(caption = "The value in the cell is the number of effect sizes") + 
   theme(plot.caption = element_text(size = 10, color = "gray10", face = "italic"),
         axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text = element_text(size = 12),
         axis.title = element_text(face = "bold", size = 12))

Box1_map3
```

## Quantitative Synthesis

```{r, echo=FALSE, out.height=600, out.width=1000, fig.width=10, fig.height=7}
est_dat <- est_dat %>%
  filter(n_studies > 1)

# Plot
ggplot(est_dat, aes(x = chemical_id, y = estimate, ymin = ci.lb, ymax = ci.ub)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.5) +
  geom_point(aes(color = pval < 0.05), size = 2.5) +
  geom_errorbar(aes(color = pval < 0.05), width = 0.2) +
  scale_color_manual(values = c("FALSE" = "grey", "TRUE" = "black"), guide = "none") +
  facet_wrap(~ health_outcome_group, scales = "free_y") +
  labs(title = "Estimated Interaction Effects",
       x = "PFAS",
       y = "Log Odds Ratio (logOR)") +
  theme_minimal(base_size = 12) +
  labs(caption = "The plot displays only meta-analytic estimates that were pooled from at least two meta-analyses. Black estimates were statistically different from 0.") + 
  theme(
    plot.caption = element_text(size = 10, color = "gray10", face = "italic"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    strip.text = element_text(size = 10, face = "bold"),
    legend.position = "none"
  )
```

## Summary Results

**1)**  Five health outcome groups out of eleven did not have more than one meta-analytic estimate for any chemical compound.  

**2)**  

**3)**  

**4)**  

# Fourth Research Question

**Question:** How consistent are the findings regarding PFAS-health outcome associations across different meta-analyses?

```{r}

```

